Udacity Project 4 EDA by STEVEN_ROSENFIELD
========================================================

> **Tip**: You will see quoted sections like this throughout the template to
help you construct your report. Make sure that you remove these notes before
you finish and submit your project!

> **Tip**: One of the requirements of this project is that your code follows
good formatting techniques, including limiting your lines to 80 characters or
less. If you're using RStudio, go into Preferences \> Code \> Display to set up
a margin line to help you keep track of this guideline!

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(ggplot2)
library(knitr)
library(dplyr)
library(grid)
library(gridExtra)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
wine <- read.csv('wineQualityReds.csv')
```

### Data Description
This data contains 1599 instances of red wine of the
Portugeses "Vinho Verde" variety.  For each instance, there are 11 variables
contain info about the chemical properties of the wine, and 1 rating that 
corresponds to the quality of the wine.  The quality of the wine is based on 
the average of 3 wine expert ratings, from 0 (very bad) to 10 (very excellent).

# Univariate Plots Section

> **Tip**: In this section, you should perform some preliminary exploration of
your dataset. Run some summaries of the data and create univariate plots to
understand the structure of the individual variables in your dataset. Don't
forget to add a comment after each plot or closely-related group of plots!
There should be multiple code chunks and text sections; the first one below is
just to help you get started.

```{r echo=FALSE, Univariate_Plots}
#run a summary of the data
str(wine)
summary(wine)
```

```{r echo=FALSE, Univariate_Plots}
#get a boxplot of all columns in order to visualize them
par(mfrow=c(3,4))
for (i in 2:length(wine)) {
        boxplot(wine[,i], main=names(wine[i]), type="l")
}
```

Some observations of these box plots:
1 - The variables with relatively high spread include citric acid, alcohol, and
quality.
2 - The variables with relatively low spread include residual sugar, chlorides,
and sulphates.
3 - The variables with approximately normal distributions include citric acid,
density, pH, alcohol, and quality.

Let's examine the dependent variable quality a little more closely.

```{r echo=FALSE, Univariate_Plots}
#histogram of quality
ggplot(data = wine, aes(x = quality)) +
  geom_histogram(binwidth = .5, color = 'black', fill = '#59A722') +
  scale_x_continuous(breaks = 1:10)
```

This plot shows a histogram of quality.  It's interesting that there were no
values less than 3, or greater than 8.  It also seems like most of the values
for quality were either 5 or 6.

```{r echo=FALSE, Univariate_Plots}
#typical is a new dataframe to show only wine observations with quality = [5,6]
typical <- subset(wine, quality == 5 | quality == 6)

#calculate the number of observations with quality = 5 or 6, divided by the \
#total number of observations
nrow(typical) / nrow(wine)
```

About 83.5% of all observations had quality = 5 or 6. This confirms my 
observation from the histogram.

My initial theory is alcohol content will have an effect on quality, so I'm
going to break down the alcohol variable by quartile, and store this data
in a new variable.

```{r echo=FALSE, Univariate_Plots}
#break down the alcohol content by quartile and store that in a new variable
cond <- wine$alcohol < 9.5
wine$alcohol.quartile <- ifelse(cond, 'low', 'mid-low')
cond <- wine$alcohol >= 10.2 & wine$alcohol < 11.1
wine$alcohol.quartile <- ifelse(cond, 'mid-high', wine$alcohol.quartile)
cond <- wine$alcohol >= 11.1
wine$alcohol.quartile <- ifelse(cond, 'high', wine$alcohol.quartile)
table(wine$alcohol.quartile)
```

```{r echo=FALSE, Univariate_Plots}
#break down the volatile acidity into halves and store that in a new variable
cond <- wine$volatile.acidity < .52
wine$vol.acid.half <- ifelse(cond, 'low', 'high')
table(wine$vol.acid.half)

wine$vol.acid.half.fac <- factor(wine$vol.acid.half, levels = c("low", "high"))
```


```{r echo=FALSE, Univariate_Plots}
#create a factor variable to organize alcohol quartile  in orderfor a histogram
wine$alc.qua.fac <- factor(wine$alcohol.quartile, 
                           levels = c("low", "mid-low","mid-high", "high"))

ggplot(wine, aes(x = wine$alc.qua.fac)) +
  geom_bar(color = 'black', fill = '#59A722')

```


# Univariate Analysis

> **Tip**: Now that you've completed your univariate explorations, it's time to
reflect on and summarize what you've found. Use the questions below to help you
gather your observations and add your own if you have other thoughts!

### What is the structure of your dataset?
The dataset includes 1599 observations of 13 variables.  All 13 variables are 
are of the type number except X and quality, which are integers.

### What is/are the main feature(s) of interest in your dataset?
The main feature of interest in my dataset is the dependent variable quality.  

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?
As of now, any of the 11 independent variables can help support my 
investigation into my feature of interest.  I will have to explore them more
in the bivariate section.

### Did you create any new variables from existing variables in the dataset?
Yes, I created the alcohol.quartile variable based on the quartile 
that the alcohol variable falls into for each observation.

I also created wine$alc.qua.fac to organize alcohol.quartile in order.

I also created a new variable to classify volatile acidity into buckets.

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?

Both residual sugar and chlorides had very long tails skewed to the right.
I did not change the form of the data, though.


# Bivariate Plots Section

> **Tip**: Based on what you saw in the univariate plots, what relationships
between variables might be interesting to look at in this section? Don't limit
yourself to relationships between a main output feature and one of the
supporting variables. Try to look at relationships between supporting variables
as well.
```{r}
str(wine)
```

```{r echo=FALSE, Bivariate_Plots}
#create a box plot of alcohol quartile vs. quality
qplot(x = alc.qua.fac, y = quality,
      data = wine,
      geom = 'boxplot')
```

Based on these box plots, it looks like alcohol is positively correlated with
quality.  Let's investigate further with a scatterplot.

```{r echo=FALSE, Bivariate_Plots}
ggplot(aes(x = alcohol, y = quality), data = wine) + 
  geom_point(alpha = 1/20, position = position_jitter(w = 0)) +
  geom_smooth(method = 'lm', color = 'red')
```

This scatter-plot shows a positive correlation between alcohol content and
quality, affirming my suspicion.

Let's see how the rest of the variables correlate with quality.

```{r echo=FALSE, Bivariate_Plots}
p1 <- ggplot(aes(x = fixed.acidity, y = quality), data = wine) + 
  geom_point(alpha = 1/20, position = position_jitter(w = 0)) +
  geom_smooth(method = 'lm', color = 'red')
p2 <- ggplot(aes(x = volatile.acidity, y = quality), data = wine) + 
  geom_point(alpha = 1/20, position = position_jitter(w = 0)) +
  geom_smooth(method = 'lm', color = 'red')
p3 <- ggplot(aes(x = citric.acid, y = quality), data = wine) + 
  geom_point(alpha = 1/20, position = position_jitter(w = 0)) +
  geom_smooth(method = 'lm', color = 'red')
p4 <- ggplot(aes(x = residual.sugar, y = quality), data = wine) + 
  geom_point(alpha = 1/20, position = position_jitter(w = 0)) +
  geom_smooth(method = 'lm', color = 'red')
p5 <- ggplot(aes(x = chlorides, y = quality), data = wine) + 
  geom_point(alpha = 1/20, position = position_jitter(w = 0)) +
  geom_smooth(method = 'lm', color = 'red')
p6 <- ggplot(aes(x = free.sulfur.dioxide, y = quality), data = wine) + 
  geom_point(alpha = 1/20, position = position_jitter(w = 0)) +
  geom_smooth(method = 'lm', color = 'red')
p7 <- ggplot(aes(x = total.sulfur.dioxide, y = quality), data = wine) + 
  geom_point(alpha = 1/20, position = position_jitter(w = 0)) +
  geom_smooth(method = 'lm', color = 'red')
p8 <- ggplot(aes(x = density, y = quality), data = wine) + 
  geom_point(alpha = 1/20, position = position_jitter(w = 0)) +
  geom_smooth(method = 'lm', color = 'red')
p9 <- ggplot(aes(x = pH, y = quality), data = wine) + 
  geom_point(alpha = 1/20, position = position_jitter(w = 0)) +
  geom_smooth(method = 'lm', color = 'red')
p10 <- ggplot(aes(x = sulphates, y = quality), data = wine) + 
  geom_point(alpha = 1/20, position = position_jitter(w = 0)) +
  geom_smooth(method = 'lm', color = 'red')
p11 <- ggplot(aes(x = alcohol, y = quality), data = wine) + 
  geom_point(alpha = 1/20, position = position_jitter(w = 0)) +
  geom_smooth(method = 'lm', color = 'red')

grid.arrange(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10, p11, ncol=4)
```

Based on these plots, it looks like alcohol, sulphates, and citric acid have
relatively large positive correlation with quality.  Also, it looks like
volatile acidity, chlorides, and total sulfur dioxide have relatively large
negative correlation with quality.

Let's calculate the Pearson correlation coefficient between all numeric
variables

```{r echo=FALSE, Bivariate_Plots}
#calculate the Pearson correlation coefficient between all numeric variables
cor(wine[2:13])
```

Based on these correlation coefficients, it looks like alcohol has the highest
correlation with quality with a Pearson corr coef = .476.  The next highest is 
sulphates with R = .251 and then citric acid with R=.226.

The most negative correlation coefficients are volatile acidity with R = -.391 
and desnity with R = -.175.  

This confirms what what I saw in the earlier graphs.

The strongest relationship is fixed acidity to citric acid with 
R = .672, and the most negative relationship is fixed acidity to pH with 
R = -.683.  These variables would logically have strong relationships because
they all describe acidity.

The highest R value for a relationship that stuck out to me was density to 
fixed acidity with R = .668.  Let's explore that more.  

```{r echo=FALSE, Bivariate_Plots}
ggplot(aes(x = density, y = fixed.acidity), data = wine) + 
  geom_point(alpha = 1/10) +
  geom_smooth(method = 'lm', color = 'blue')
```

High correlation in this graph!

Now I want to explore volatile acidity's correlation with quality, because
I noted that it has a strong negative correlation.

```{r echo=FALSE, Bivariate_Plots}
#create a box plot of volatile acidity half vs. quality
qplot(x = vol.acid.half.fac, y = quality,
      data = wine,
      geom = 'boxplot')
```

This isn't a very useful graph, but it does show that the volatile acidity 
causes the quality to have a higher median when separated into buckets.

Let's group the quality into groups and analyze the data that way

```{r}
#calculate the alcohol mean and median by quality bucket
quality_groups <- group_by(wine,quality)
wine.alc_by_qual <- summarise(quality_groups,
                             alc_mean = mean(alcohol),
                             alc_median = median(alcohol),
                             n = n())

wine.alc_by_qual <- arrange(wine.alc_by_qual,quality)

head(wine.alc_by_qual)
```

```{r}
ggplot(aes(quality, alc_mean), data = wine.alc_by_qual) + 
  geom_line(color = 'purple')
```

This graph shows a strong correlation between wine quality and the mean value 
of alcohol content for each quality.

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

Based on these correlation coefficients, it looks like alcohol has the highest
correlation with quality (Pearson corr coef = .476).  The next highest is 
sulphates with R = .251 and then citric acid with R=.226.

The most negative correlation coefficients with quality 
are volatile acidity with R = -.391 and density with R = -.175.  

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

The strongest relationship out of all the data is fixed acidity to citric acid 
with R = .672, and the most negative relationship is fixed acidity to pH with 
R = -.683.  These variables would logically have strong relationships because
they all describe acidity.

The highest R value for a relationship that stuck out to me was density to 
fixed acidity with R = .668. 

### What was the strongest relationship you found?

As mentioned earlier, the strongest relationship I saw was fixed acidity to
citric acid with R = .672.  However, this makes sense because they both
describe acidity.

The highest R value for two variables that didn't definitely describe the same
thing was density to fixed acidity with R = .668. 

# Multivariate Plots Section

> **Tip**: Now it's time to put everything together. Based on what you found in
the bivariate plots section, create a few multivariate plots to investigate
more complex interactions between variables. Make sure that the plots that you
create here are justified by the plots you explored in the previous section. If
you plan on creating any mathematical models, this is the section where you
will do that.

```{r echo=FALSE, Multivariate_Plots}

```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

------

# Final Plots and Summary

> **Tip**: You've done a lot of exploration and have built up an understanding
of the structure of and relationships between the variables in your dataset.
Here, you will select three plots from all of your previous exploration to
present here as a summary of some of your most interesting findings. Make sure
that you have refined your selected plots for good titling, axis labels (with
units), and good aesthetic choices (e.g. color, transparency). After each plot,
make sure you justify why you chose each plot by describing what it shows.

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection

> **Tip**: Here's the final step! Reflect on the exploration you performed and
the insights you found. What were some of the struggles that you went through?
What went well? What was surprising? Make sure you include an insight into
future work that could be done with the dataset.

> **Tip**: Don't forget to remove this, and the other **Tip** sections before
saving your final work and knitting the final report!