Udacity Project 4 EDA by STEVEN_ROSENFIELD
========================================================

> **Tip**: You will see quoted sections like this throughout the template to
help you construct your report. Make sure that you remove these notes before
you finish and submit your project!

> **Tip**: One of the requirements of this project is that your code follows
good formatting techniques, including limiting your lines to 80 characters or
less. If you're using RStudio, go into Preferences \> Code \> Display to set up
a margin line to help you keep track of this guideline!

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(ggplot2)
library(knitr)
library(dplyr)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
wine <- read.csv('wineQualityReds.csv')
```

### Data Description
This data contains 1599 instances of red wine of the
Portugeses "Vinho Verde" variety.  For each instance, there are 11 variables
contain info about the chemical properties of the wine, and 1 rating that 
corresponds to the quality of the wine.  The quality of the wine is based on 
the average of 3 wine expert ratings, from 0 (very bad) to 10 (very excellent).

# Univariate Plots Section

> **Tip**: In this section, you should perform some preliminary exploration of
your dataset. Run some summaries of the data and create univariate plots to
understand the structure of the individual variables in your dataset. Don't
forget to add a comment after each plot or closely-related group of plots!
There should be multiple code chunks and text sections; the first one below is
just to help you get started.

```{r echo=FALSE, Univariate_Plots}
#run a summary of the data
str(wine)
summary(wine)
```

```{r echo=FALSE, Univariate_Plots}
#get a boxplot of all columns in order to visualize them
par(mfrow=c(3,4))
for (i in 2:length(wine)) {
        boxplot(wine[,i], main=names(wine[i]), type="l")
}
```

Some observations of these box plots:
1 - The variables with relatively high spread include citric acid, alcohol, and
quality.
2 - The variables with relatively low spread include residual sugar, chlorides,
and sulphates.
3 - The variables with approximately normal distributions include citric acid,
density, pH, alcohol, and quality.

Let's examine the dependent variable quality a little more closely.

```{r echo=FALSE, Univariate_Plots}
#histogram of quality
ggplot(data = wine, aes(x = quality)) +
  geom_histogram(binwidth = .5, color = 'black', fill = '#59A722') +
  scale_x_continuous(breaks = 1:10)
```

This plot shows a histogram of quality.  It's interesting that there were no
values less than 3, or greater than 8.  It also seems like most of the values
for quality were either 5 or 6.

```{r echo=FALSE, Univariate_Plots}
#typical is a new dataframe to show only wine observations with quality = [5,6]
typical <- subset(wine, quality == 5 | quality == 6)

#calculate the number of observations with quality = 5 or 6, divided by the \
#total number of observations
nrow(typical) / nrow(wine)
```

About 83.5% of all observations had quality = 5 or 6. This confirms my 
observation from the histogram.

My initial theory is alcohol content will have an effect on quality, so I'm
going to break down the alcohol variable by quartile, and store this data
in a new variable.

```{r echo=FALSE, Univariate_Plots}
#break down the alcohol content by quartile and store that in a new variable
cond <- wine$alcohol < 9.5
wine$alcohol.quartile <- ifelse(cond, 'low', 'mid-low')
cond <- wine$alcohol >= 10.2 & wine$alcohol < 11.1
wine$alcohol.quartile <- ifelse(cond, 'mid-high', wine$alcohol.quartile)
cond <- wine$alcohol >= 11.1
wine$alcohol.quartile <- ifelse(cond, 'high', wine$alcohol.quartile)
table(wine$alcohol.quartile)
```


> **Tip**: Make sure that you leave a blank line between the start / end of
each code block and the end / start of your Markdown text so that it is
formatted nicely in the knitted text. Note as well that text on consecutive
lines is treated as a single space. Make sure you have a blank line between
your paragraphs so that they too are formatted for easy readability.

# Univariate Analysis

> **Tip**: Now that you've completed your univariate explorations, it's time to
reflect on and summarize what you've found. Use the questions below to help you
gather your observations and add your own if you have other thoughts!

### What is the structure of your dataset?
The dataset includes 1599 observations of 13 variables.  All 13 variables are 
are of the type number except X and quality, which are integers.

### What is/are the main feature(s) of interest in your dataset?
The main feature of interest in my dataset is the dependent variable quality.  

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?
As of now, any of the 11 independent variables can help support my 
investigation into my feature of interest.  I will have to explore them more
in the bivariate section.

### Did you create any new variables from existing variables in the dataset?
Yes, I created the alcohol.quartile variable based on the quartile 
that the alcohol variable falls into for each observation.

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?

Both residual sugar and chlorides had very long tails skewed to the right.
I did not change the form of the data, though.


# Bivariate Plots Section

> **Tip**: Based on what you saw in the univariate plots, what relationships
between variables might be interesting to look at in this section? Don't limit
yourself to relationships between a main output feature and one of the
supporting variables. Try to look at relationships between supporting variables
as well.

```{r echo=FALSE, Bivariate_Plots}

```

# Bivariate Analysis

> **Tip**: As before, summarize what you found in your bivariate explorations
here. Use the questions below to guide your discussion.

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

### What was the strongest relationship you found?


# Multivariate Plots Section

> **Tip**: Now it's time to put everything together. Based on what you found in
the bivariate plots section, create a few multivariate plots to investigate
more complex interactions between variables. Make sure that the plots that you
create here are justified by the plots you explored in the previous section. If
you plan on creating any mathematical models, this is the section where you
will do that.

```{r echo=FALSE, Multivariate_Plots}

```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

------

# Final Plots and Summary

> **Tip**: You've done a lot of exploration and have built up an understanding
of the structure of and relationships between the variables in your dataset.
Here, you will select three plots from all of your previous exploration to
present here as a summary of some of your most interesting findings. Make sure
that you have refined your selected plots for good titling, axis labels (with
units), and good aesthetic choices (e.g. color, transparency). After each plot,
make sure you justify why you chose each plot by describing what it shows.

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection

> **Tip**: Here's the final step! Reflect on the exploration you performed and
the insights you found. What were some of the struggles that you went through?
What went well? What was surprising? Make sure you include an insight into
future work that could be done with the dataset.

> **Tip**: Don't forget to remove this, and the other **Tip** sections before
saving your final work and knitting the final report!