Udacity Project 4 EDA by STEVEN_ROSENFIELD
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(ggplot2)
library(knitr)
library(dplyr)
library(grid)
library(gridExtra)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Load_the_Data}
# Load the Data
wine <- read.csv('wineQualityReds.csv')
```

### Data Description
This data contains 1599 instances of red wine of the
Portugeses "Vinho Verde" variety.  For each instance, there are 11 variables
contain info about the chemical properties of the wine, and 1 rating that 
corresponds to the quality of the wine.  The quality of the wine is based on 
the average of 3 wine expert ratings, from 0 (very bad) to 10 (very excellent).

# Univariate Plots Section

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots1}
#run a summary of the data
str(wine)
summary(wine)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots2}
#get a boxplot of all columns in order to visualize them
par(mfrow=c(3,4))
for (i in 2:length(wine)) {
        boxplot(wine[,i], main=names(wine[i]), type="l")
}
```

Some observations of these box plots:
1 - The variables with relatively high spread include citric acid, alcohol, and
quality.
2 - The variables with relatively low spread include residual sugar, chlorides,
and sulphates.
3 - The variables with approximately normal distributions include citric acid,
density, pH, alcohol, and quality.

Let's examine the dependent variable quality a little more closely.

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots3}
#histogram of quality
ggplot(data = wine, aes(x = quality)) +
  geom_histogram(binwidth = .5, color = 'black', fill = '#59A722') +
  scale_x_continuous(breaks = 1:10)
```

This plot shows a histogram of quality.  It's interesting that there were no
values less than 3, or greater than 8.  It also seems like most of the values
for quality were either 5 or 6.

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots4}
#typical is a new dataframe to show only wine observations with quality = [5,6]
typical <- subset(wine, quality == 5 | quality == 6)

#calculate the number of observations with quality = 5 or 6, divided by the \
#total number of observations
nrow(typical) / nrow(wine)
```

About 83.5% of all observations had quality = 5 or 6. This confirms my 
observation from the histogram.

My initial theory is alcohol content will have an effect on quality, so I'm
going to break down the alcohol variable by quartile, and store this data
in a new variable called alcohol.quartile with values of 'low', 'mid-low', 
'mid-high', and 'high' depending on the quartile the alcohol content falls into.

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots5}
#break down the alcohol content by quartile and store that in a new variable
cond <- wine$alcohol < 9.5
wine$alcohol.quartile <- ifelse(cond, 'low', 'mid-low')
cond <- wine$alcohol >= 10.2 & wine$alcohol < 11.1
wine$alcohol.quartile <- ifelse(cond, 'mid-high', wine$alcohol.quartile)
cond <- wine$alcohol >= 11.1
wine$alcohol.quartile <- ifelse(cond, 'high', wine$alcohol.quartile)
table(wine$alcohol.quartile)
```

Now I will break down the volatile acidity into halves (above median vs.
below median) and label each row with'low' or 'high' and store this data
ina new variable vol.acid.half.  I will also create a factor variable for
volatile acidity quartile to order it correctly for plotting.


```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots6}
#break down the volatile acidity into halves and store that in a new variable
cond <- wine$volatile.acidity < .52
wine$vol.acid.half <- ifelse(cond, 'low', 'high')
table(wine$vol.acid.half)

wine$vol.acid.half.fac <- factor(wine$vol.acid.half, levels = c("low", "high"))
```

Now I will create a factor variable for alcohol quartile to order it correctly
for plotting.

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots7}
#create a factor variable to organize alcohol quartile  in orderfor a histogram
wine$alc.qua.fac <- factor(wine$alcohol.quartile, 
                           levels = c("low", "mid-low","mid-high", "high"))

ggplot(wine, aes(x = wine$alc.qua.fac)) +
  geom_bar(color = 'black', fill = '#59A722')

```

It looks like the buckets are close to evenly distributed.

Let's break the sulphates into buckets as well by rounding each value down to
the nearest tenth and storing the result in a variable called 
sulphates_bucket.

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots8}
#get sulphates buckets
wine$sulphates_bucket <- (floor(wine$sulphates*10))/10
table(wine$sulphates_bucket)

```

There is a low amount of wine with sulphates>=1, so let's group them all into
one bucket.  Also, lets group the .3's in with the .4's because there are only
9 wines with a value of .3.

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots9}
cond <- wine$sulphates_bucket >= 1
wine$sulphates_bucket <- ifelse(cond, 1, wine$sulphates_bucket)
cond <- wine$sulphates_bucket ==.3
wine$sulphates_bucket <- ifelse(cond, .4, wine$sulphates_bucket)
table(wine$sulphates_bucket)

```

This is the number of occurences for each sulphates bucket.
It looks like cleaner data to work with!

Let's create a histogram to confirm.

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots10}
ggplot(wine, aes(x = wine$sulphates_bucket)) +
  geom_bar(color = 'black', fill = '#59A722')

```

The sulphates buckets look approximately normally distributed.

# Univariate Analysis

### What is the structure of your dataset?
The dataset includes 1599 observations of 13 variables.  All 13 variables are 
are of the type number except X and quality, which are integers.

### What is/are the main feature(s) of interest in your dataset?
The main feature of interest in my dataset is the dependent variable quality.  

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?
As of now, any of the 11 independent variables can help support my 
investigation into my feature of interest.  I will have to explore them more
in the bivariate section.

### Did you create any new variables from existing variables in the dataset?
Yes, I created the alcohol.quartile variable based on the quartile 
that the alcohol variable falls into for each observation.

I also created wine$alc.qua.fac to organize alcohol.quartile in order.

I also created a new variable to classify volatile acidity into buckets.

I also created a new variable to classify sulphates into buckets.

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?

Both residual sugar and chlorides had very long tails skewed to the right.

Sulphates also had a relatively long tail skewed right, so when I grouped 
sulphates into buckets with a new variable, I grouped the outliers into 
buckets that will make the data easier to visualize going forward.


# Bivariate Plots Section

Let's create a box plot of alcohol vs. quality

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots1}
#create a box plot of alcohol quartile vs. quality
qplot(x = alc.qua.fac, y = quality,
      data = wine,
      geom = 'boxplot')
```

Based on these box plots, it looks like alcohol is positively correlated with
quality.  Let's investigate further with a scatterplot.

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots2}
ggplot(aes(x = alcohol, y = quality), data = wine) + 
  geom_point(alpha = 1/20, position = position_jitter(w = 0)) +
  geom_smooth(method = 'lm', color = 'red')
```

This scatter-plot shows a positive correlation between alcohol content and
quality, affirming my suspicion.

Let's see how the rest of the variables correlate with quality.

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots3}
p1 <- ggplot(aes(x = fixed.acidity, y = quality), data = wine) + 
  geom_point(alpha = 1/20, position = position_jitter(w = 0)) +
  geom_smooth(method = 'lm', color = 'red')
p2 <- ggplot(aes(x = volatile.acidity, y = quality), data = wine) + 
  geom_point(alpha = 1/20, position = position_jitter(w = 0)) +
  geom_smooth(method = 'lm', color = 'red')
p3 <- ggplot(aes(x = citric.acid, y = quality), data = wine) + 
  geom_point(alpha = 1/20, position = position_jitter(w = 0)) +
  geom_smooth(method = 'lm', color = 'red')
p4 <- ggplot(aes(x = residual.sugar, y = quality), data = wine) + 
  geom_point(alpha = 1/20, position = position_jitter(w = 0)) +
  geom_smooth(method = 'lm', color = 'red')
p5 <- ggplot(aes(x = chlorides, y = quality), data = wine) + 
  geom_point(alpha = 1/20, position = position_jitter(w = 0)) +
  geom_smooth(method = 'lm', color = 'red')
p6 <- ggplot(aes(x = free.sulfur.dioxide, y = quality), data = wine) + 
  geom_point(alpha = 1/20, position = position_jitter(w = 0)) +
  geom_smooth(method = 'lm', color = 'red')
p7 <- ggplot(aes(x = total.sulfur.dioxide, y = quality), data = wine) + 
  geom_point(alpha = 1/20, position = position_jitter(w = 0)) +
  geom_smooth(method = 'lm', color = 'red')
p8 <- ggplot(aes(x = density, y = quality), data = wine) + 
  geom_point(alpha = 1/20, position = position_jitter(w = 0)) +
  geom_smooth(method = 'lm', color = 'red')
p9 <- ggplot(aes(x = pH, y = quality), data = wine) + 
  geom_point(alpha = 1/20, position = position_jitter(w = 0)) +
  geom_smooth(method = 'lm', color = 'red')
p10 <- ggplot(aes(x = sulphates, y = quality), data = wine) + 
  geom_point(alpha = 1/20, position = position_jitter(w = 0)) +
  geom_smooth(method = 'lm', color = 'red')
p11 <- ggplot(aes(x = alcohol, y = quality), data = wine) + 
  geom_point(alpha = 1/20, position = position_jitter(w = 0)) +
  geom_smooth(method = 'lm', color = 'red')

grid.arrange(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10, p11, ncol=4)
```

Based on these plots, it looks like alcohol, sulphates, and citric acid have
relatively large positive correlation with quality.  Also, it looks like
volatile acidity, chlorides, and total sulfur dioxide have relatively large
negative correlation with quality.

Let's calculate the Pearson correlation coefficient between all numeric
variables

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots4}
#calculate the Pearson correlation coefficient between all numeric variables
cor(wine[2:13])
```

Based on these correlation coefficients, it looks like alcohol has the highest
correlation with quality with a Pearson corr coef = .476.  The next highest is 
sulphates with R = .251 and then citric acid with R=.226.

The most negative correlation coefficients are volatile acidity with R = -.391 
and density with R = -.175.  

This confirms what what I saw in the earlier graphs.

The strongest relationship is fixed acidity to citric acid with 
R = .672, and the most negative relationship is fixed acidity to pH with 
R = -.683.  These variables would logically have strong relationships because
they all describe acidity.

The highest R value for a relationship that stuck out to me was density to 
fixed acidity with R = .668.  Let's explore that more.  

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots5}
ggplot(aes(x = density, y = fixed.acidity), data = wine) + 
  geom_point(alpha = 1/10) +
  geom_smooth(method = 'lm', color = 'blue')
```

High correlation in this graph!

Let's look at alcohol vs. volatile acidity because it looks like these two
variables have the strongest correlations to quality

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots6}
ggplot(aes(x = alcohol, y = volatile.acidity), data = wine) + 
  geom_point(alpha = 1/10, color = 'darkblue') +
  geom_smooth(method = 'lm', color = 'yellow')
```

There's a slight negative correlation, which makes sense because they had
a correlation coefficient = -.202

Now I want to explore volatile acidity's correlation with quality, because
I noted that it has a strong negative correlation.

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots7}
#create a box plot of volatile acidity half vs. quality
qplot(x = vol.acid.half.fac, y = quality,
      data = wine,
      geom = 'boxplot')
```

This isn't a very useful graph, but it does show that the volatile acidity 
causes the quality to have a higher median when separated into buckets.

Now I want to explore sulphates's correlation with quality, because
I noted that it has a relatively high positive correlation.

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots8}
#create a box plot of volatile acidity half vs. quality
qplot(x = sulphates_bucket, y = quality,
      data = wine,
      geom = 'boxplot')
```

The slight correlation clearly does not appear in this chart, as it grouped
all buckets into a single boxplot.

Let's group the quality into groups and analyze the data that way.

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots9}
#calculate the alcohol mean and median by quality bucket
quality_groups <- group_by(wine,quality)
wine.alc_by_qual <- summarise(quality_groups,
                             alc_mean = mean(alcohol),
                             alc_median = median(alcohol),
                             n = n())

wine.alc_by_qual <- arrange(wine.alc_by_qual,quality)

head(wine.alc_by_qual)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots10}
ggplot(aes(quality, alc_mean), data = wine.alc_by_qual) + 
  geom_line(color = 'purple')
```

This graph shows a strong correlation between wine quality and the mean value 
of alcohol content for each quality.

Let's look at the strongest negative correlation, volatile acidity,
compared to quality groups now.

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots11}
#calculate the volatile acidity mean and median by quality bucket
wine.vol_acid_by_qual <- summarise(quality_groups,
                             vol_acid_mean = mean(volatile.acidity),
                             vol_acid_median = median(volatile.acidity),
                             n = n())

wine.vol_acid_by_qual <- arrange(wine.vol_acid_by_qual,quality)

head(wine.vol_acid_by_qual)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots12}
ggplot(aes(quality, vol_acid_mean), data = wine.vol_acid_by_qual) + 
  geom_line(color = 'purple')
```

This graph shows a strong negative correlation between wine quality and the 
mean value of volatile acidity for each quality.

Let's look at the sulphates compared to quality groups now.

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots13}
#calculate the volatile acidity mean and median by quality bucket
wine.sulphates_by_qual <- summarise(quality_groups,
                             sulphates_mean = mean(sulphates),
                             sulphates_median = median(sulphates),
                             n = n())

wine.sulphates_by_qual <- arrange(wine.sulphates_by_qual,quality)

head(wine.vol_acid_by_qual)


ggplot(aes(quality, sulphates_mean), data = wine.sulphates_by_qual) + 
  geom_line(color = 'purple')
```

There is clearly a positive correlation between quality and sulphates.

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

Based on these correlation coefficients, it looks like alcohol has the highest
correlation with quality (Pearson corr coef = .476).  The next highest is 
sulphates with R = .251 and then citric acid with R=.226.

The most negative correlation coefficients with quality 
are volatile acidity with R = -.391 and density with R = -.175.  

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

The strongest relationship out of all the data is fixed acidity to citric acid 
with R = .672, and the most negative relationship is fixed acidity to pH with 
R = -.683.  These variables would logically have strong relationships because
they all describe acidity.

The highest R value for a relationship that stuck out to me was density to 
fixed acidity with R = .668. 

### What was the strongest relationship you found?

As mentioned earlier, the strongest relationship I saw was fixed acidity to
citric acid with R = .672.  However, this makes sense because they both
describe acidity.

The highest R value for two variables that didn't definitely describe the same
thing was density to fixed acidity with R = .668. 

# Multivariate Plots Section

Let's look at the two strongest correlated variables with quality in one graph,
alcohol and volatile acidity.

```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_Plots1}
ggplot(aes(x = volatile.acidity, y = quality), data = wine) + 
  geom_line(aes(color = alc.qua.fac), stat = 'summary', fun.y = median)
```

This chart isn't very useful because it's too busy.  Let's try a scatterplot.

```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_Plots2}
ggplot(aes(x = volatile.acidity, y = quality),
       data = wine) +
  geom_point(aes(color = alc.qua.fac), alpha = 1/3)
```

It looks like what I expected. The higher quality wine is more purple (high
alcohol content), and falls towards the left on the graph (low volatile
acidity).

Now let's try setting alcohol as the x-axis and volatile acidity as the 
categorical variable.

```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_Plots3}
ggplot(aes(x = alcohol, y = quality),
       data = wine) +
  geom_point(aes(color = vol.acid.half.fac), alpha = 1/3)
```

This is a good looking chart!  It clearly shows as alcohol content increases
quality increases, and lower volatile acidity corresponds to higher quality.

Let's look at alcohol vs. volatile acidity broken into sulphate groups.  These
are the 3 most correlated variables to quality

```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_Plots4}
ggplot(aes(x = alcohol, y = volatile.acidity), data = wine) +
  geom_point(aes(color = sulphates_bucket), alpha = 1/4)
```

This graph doesn't seem to imply much correlation between these variables at
all.

Let's look at sulphates bucket vs. quality broken down by volatile acidity.

```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_Plots5}
ggplot(aes(x = alcohol, y = quality),
       data = wine) +
  geom_point(aes(color = sulphates_bucket), alpha = 1/3)
```

Again this isn't a very useful chart, because it doesn't show much correlation.

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

I observed that alcohol content corresponds to increased quality, 
whereas volatile acidity content corresponds to decreased quality.

Sulphates seemed arbritrary with respect to quality.

### Were there any interesting or surprising interactions between features?

The most interesting interaction between features is that alcohol content
corresponds to increased quality, whereas volatile acidity content corresponds
to decreased quality.

Other than that, it wasn't too interesting because there weren't many
attributes that had a large affect on quality.

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

I did not create any models with my dataset.

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_One}
p1 <- ggplot(aes(x = fixed.acidity, y = quality), data = wine) + 
  geom_point(alpha = 1/20, position = position_jitter(w = 0)) +
  geom_smooth(method = 'lm', color = 'red') +
  ggtitle('Fixed Acidity vs. Quality') +
  labs(x = "fixed acidity (tartaric acid - g / dm^3)")+ 
  theme(plot.title = element_text(size = 8, face = "bold"), 
        axis.title =element_text(size=6)) 
p2 <- ggplot(aes(x = volatile.acidity, y = quality), data = wine) + 
  geom_point(alpha = 1/20, position = position_jitter(w = 0)) +
  geom_smooth(method = 'lm', color = 'red') +
  ggtitle('Volatile Acidity vs. Quality') +
  labs(x = "volatile acidity (acetic acid - g / dm^3)")+ 
  theme(plot.title = element_text(size = 8, face = "bold"), 
        axis.title =element_text(size=6)) 
p3 <- ggplot(aes(x = citric.acid, y = quality), data = wine) + 
  geom_point(alpha = 1/20, position = position_jitter(w = 0)) +
  geom_smooth(method = 'lm', color = 'red') +
  ggtitle('Citric Acid vs. Quality') +
  labs(x = "citric acid (g / dm^3)")+ 
  theme(plot.title = element_text(size = 8, face = "bold"), 
        axis.title =element_text(size=6)) 
p4 <- ggplot(aes(x = residual.sugar, y = quality), data = wine) + 
  geom_point(alpha = 1/20, position = position_jitter(w = 0)) +
  geom_smooth(method = 'lm', color = 'red') +
  ggtitle('Residual Sugar vs. Quality') +
  labs(x = "residual sugar (g / dm^3)")+ 
  theme(plot.title = element_text(size = 8, face = "bold"), 
        axis.title =element_text(size=6)) 
p5 <- ggplot(aes(x = chlorides, y = quality), data = wine) + 
  geom_point(alpha = 1/20, position = position_jitter(w = 0)) +
  geom_smooth(method = 'lm', color = 'red') +
  ggtitle('Chlorides vs. Quality') +
  labs(x = "chlorides (sodium chloride - g / dm^3")+ 
  theme(plot.title = element_text(size = 8, face = "bold"), 
        axis.title =element_text(size=6)) 
p6 <- ggplot(aes(x = free.sulfur.dioxide, y = quality), data = wine) + 
  geom_point(alpha = 1/20, position = position_jitter(w = 0)) +
  geom_smooth(method = 'lm', color = 'red') +
  ggtitle('Free Sulfur Dioxide vs. Quality') +
  labs(x = "free sulfur dioxide (mg / dm^3)")+ 
  theme(plot.title = element_text(size = 7, face = "bold"), 
        axis.title =element_text(size=6)) 
p7 <- ggplot(aes(x = total.sulfur.dioxide, y = quality), data = wine) + 
  geom_point(alpha = 1/20, position = position_jitter(w = 0)) +
  geom_smooth(method = 'lm', color = 'red') +
  ggtitle('Total Sulfur Dioxide vs. Quality') +
  labs(x = "total sulfur dioxide (mg / dm^3)")+ 
  theme(plot.title = element_text(size = 7, face = "bold"), 
        axis.title =element_text(size=6)) 
p8 <- ggplot(aes(x = density, y = quality), data = wine) + 
  geom_point(alpha = 1/20, position = position_jitter(w = 0)) +
  geom_smooth(method = 'lm', color = 'red') +
  ggtitle('Density vs. Quality') +
  labs(x = "density (g / cm^3)")+ 
  theme(plot.title = element_text(size = 8, face = "bold"), 
        axis.title =element_text(size=6)) 
p9 <- ggplot(aes(x = pH, y = quality), data = wine) + 
  geom_point(alpha = 1/20, position = position_jitter(w = 0)) +
  geom_smooth(method = 'lm', color = 'red') +
  ggtitle('pH vs. Quality') +
  labs(x = "pH")+ 
  theme(plot.title = element_text(size = 8, face = "bold"), 
        axis.title =element_text(size=6)) 
p10 <- ggplot(aes(x = sulphates, y = quality), data = wine) + 
  geom_point(alpha = 1/20, position = position_jitter(w = 0)) +
  geom_smooth(method = 'lm', color = 'red') +
  ggtitle('Sulphates vs. Quality') +
  labs(x = "sulphates (potassium sulphate - g / dm3)")+ 
  theme(plot.title = element_text(size = 8, face = "bold"), 
        axis.title =element_text(size=6)) 
p11 <- ggplot(aes(x = alcohol, y = quality), data = wine) + 
  geom_point(alpha = 1/20, position = position_jitter(w = 0)) +
  geom_smooth(method = 'lm', color = 'red') +
  ggtitle('Alcohol vs. Quality') +
  labs(x = "alcohol (% by volume)")+ 
  theme(plot.title = element_text(size = 8, face = "bold"), 
        axis.title =element_text(size=6)) 

grid.arrange(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10, p11, ncol=4)
```

### Description One
I chose this plot because it shows how each of the attributes can affect
the variable of intereset, quality.  It helped me gather insight into the data
to know which attributes I should explore more closely, and which attributes
I can probably disregard because they have a loose connection to quality.

### Plot Two
```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_Two}
ggplot(aes(quality, alc_mean), data = wine.alc_by_qual) + 
  geom_line(color = 'purple') + 
  ggtitle('Quality Group vs. Alcohol Mean') +
  labs(x = "Quality Group (score between 0 and 10)", 
       y = "Alcohol Mean (% by volume)")+ 
  theme(plot.title = element_text(size = 24, face = "bold"), 
        axis.title =element_text(size=14)) 
```

### Description Two
Despite it's simplicity, this plot shows how the quality of wine varies
directly with alcohol content.  To make this plot, I grouped the data by 
quality score, and then took the mean alcohol content for each quality group.
This plot shows a strong correlation between quality and alcohol, and it seems
alcohol content drives the quality score more than any other attribute.

### Plot Three
```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_Three}
ggplot(aes(x = alcohol, y = quality), data = wine) +
  geom_point(aes(color = vol.acid.half.fac), alpha = 1/3) +
  ggtitle('Quality vs. Alcohol by Volatile Acidity') +
  labs(y = "Quality (score between 0 and 10)", 
       x = "Alcohol (% by volume)")+ 
  theme(plot.title = element_text(size = 24, face = "bold"), 
        axis.title =element_text(size=14)) +
  labs(colour = "Volatile Acidity Group")
```

### Description Three
I chose this plot because it shows the two attributes with the 
strongest correlation to quality.  As Alcohol % by volumne increases,
the quality increases.  On the other hand, wines in the high Volatile Acidity
group are more likely to be a lower quality wine than wines in the low Volatile
Acidity group.

------

# Reflection

This project taught me a lot about how to use R to perform Exploratory Data
Analysis.  I chose this dataset because I like to drink wine, and I was looking 
forward to seeing if there were any scientific reasons as to why one bottle of
wine was better than another bottle of wine.  I was able to gather some 
insight into what makes experts consider a bottle of wine good.  However,
the dataset did not allow me to gain as much insight as I would've liked for
several reasons.

If these 3 experts' ratings are to be trusted, there were many insights I found
into wine quality.  First of all, I learned that higher alcohol content often
contributes to better quality wine.  Also, I learned that high
volatile acidity contributes to lower quality wine.  This makes sense because
this is "the amount of acetic acid in wine, which at too high of levels can 
lead to an unpleasant, vinegar taste."  Out of the other 9 variables, I couldn't
gather too much insight into what constitutes good quality wine, because
they had low correlations to the quality variable.

Unfortunately, the dataset I chose didn't allow me to make too many interesting
insights or visualizations for several reasons.  First of all, the quality 
values were only integers.  This was very frustrating because if this value
took fractions, I could've gathered more accurate insights into how other 
variables contributed to it.  Also, the average of the 3 experts' ratings
naturally should've been decimals, so it was confusing why the data only had
integers.

Another reason the data wasn't great was there wasn't much correlation
between the variables.  For example no variable given in the dataset had
a higher correlation coefficient than .5 with quality.  
If there was better correlation, the graphs would've looked better and more 
insightful for my audience.

If future work was going to be done with this dataset, I would suggest
collecting data from more than 3 experts, and then storing the quality rating
as a decimal in order to get more granular insight into how other variables
contributed to the rating.

Overall, I got valuable experience working with R and performing Exploratory
Data Analysis, but I wish I had chosen a better dataset to work with for this
project.

Thanks for reading!

# Resources
www.stackoverflow.com

https://s3.amazonaws.com/content.udacity-data.com/courses/ud651/diamondsExample_2016-05.html
